---
title: "EEMS Pipeline"
author: "Henry Sanford"
date: '2022-05-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ape) # for calculating pairwise distances
library(TreeTools) # for reading nexus files
library(readxl)
library(phylotools)
library(rEEMSplots)
library(rgdal)
library(graphics)
library(sf)
library(ggplot2)
library(dplyr)
```

## Specifying files

```{r}
# Name of species or group, used to name input files
input_name <- "coleonyx"
# Where to store input files for the EEMS run
input_directory <- '/media/digitalstorm/Storage1/hayden/coleonyx/EEMS/input'
# Coordinates file, which must have "Catalogue", "lat" and "lon" columns
coordinates <- "/media/digitalstorm/Storage1/hayden/coleonyx/EEMS/data/geolocation.xlsx"
# Nexus file path
nexus <- "/media/digitalstorm/Storage1/hayden/coleonyx/coleonyx_noout_outfiles/coleonyx_thin.recode.min4.nexus"
```

## Creating dissimilarity matrix

```{r}
seqs <- ReadCharacters(nexus)
seqmatrix <- as.matrix(seqs)
dist <- dist.gene(seqmatrix, method = "percentage")
matrix <-as.matrix(dist)
heatmap(matrix) # visualize pairwise distances as a heatmap
```
# Create a table of coordinates with the same order as the sequence matrix
Any samples not present in the coordinates file will be dropped
```{r}
identifiers <- rownames(as.data.frame(seqs[,0]))
coords <- read_excel(coordinates)
coords <- coords[match(identifiers, coords$Catalogue),]
plot(coords[c("lat", "lon")]) # Plot to check accuracy
```

## Creating boundary box

The boundary box should be made in Google Earth Pro. 
1. Import the coordinates
2. Add > Polygon and draw a polygon around the coordinates
3. Copy and Paste the polygon to a text file
4. Edit to only include the coordinates, in the order latitude, longitude
5. Coordinates should be separated by spaces, not commas
6. Save as *your input name*.boundary to the input directory

## Create files for EEMS run
Write all the coordinates and sequence matrix to the input directory

```{r}
if (!dir.exists(input_directory)){
  dir.create(input_directory)
}
write.table(matrix, file=paste(input_directory, "/", input_name, ".diffs", sep=""), row.names=FALSE, col.names=FALSE)
write.table(coords, file=paste(input_directory, "/", input_name, ".coord", sep=""), row.names=FALSE, col.names=FALSE)
```

## Write parameters to params.ini
```{r}
mcmcpath = "path"
numBurnIter = 10000000
numMCMCIter = 50000000
numThinIter = 9999
diploid = "false"
datapath = "input/coleonyx"
nDemes = 500
nIndiv = nrow(coords)
nSites = length(seqs[1,])
parameter_names <- c("datapath","mcmcpath","nIndiv","nSites","nDemes","diploid","numMCMCIter","numBurnIter","numThinIter")
parameter_values <- c(datapath, mcmcpath, nIndiv, nSites, nDemes, diploid, numMCMCIter, numBurnIter, numThinIter)
params <- cbind(parameter_names, parameter_values)
write.table(params, col.names = FALSE, row.names = FALSE, file = "test_params.ini", sep = " = ", quote = FALSE)
```

## Run EEMS from the console

Edit params.ini to correct nIndiv, nSites
Use nSites from the nexus file
Use nIndiv from the coordinates file
Set mcmcpath to an empty directory to store results

From the terminal: 
../runeems_snps --params params.ini 

## Change coordinate projections for plotting

```{r}
projection_none <- "+proj=longlat +datum=WGS84"
projection_mercator <- "+proj=merc +datum=WGS84"
coords_longlat = coords[,c(2,1)]
plot(coords_longlat)
# Create an sf object from coordinates
sf <- st_as_sf(x = coords_longlat, coords = c("lon","lat"), crs = CRS(projection_none))
# Transform to the Mercator projection
sf_merc <- st_transform(x = sf, crs = CRS(projection_mercator))
# Extract coordinates from SF object
coords_merc <- as.matrix(st_coordinates(sf_merc))
plot(coords_merc)
```


## Plot results!

```{r}
assem_direc <- paste0(run_directory, threshold)
plot_figs <- function(chain_dir){
  figure_path <- file.path(assem_direc, "figures")
  figure_name <- paste0(figure_path, '/', assem_direc, '_', chain_dir)
  print(figure_name)
  eems_plot <- eems.plots(mcmcpath = paste(run_directory, threshold, "/out/", chain_dir, sep=""),
   plotpath = figure_name,  
   m.plot.xy = {
     points(coords_merc, col = "purple");
   },
   q.plot.xy = {
     points(coords_merc, col = "purple");
   },
   projection.in = projection_none,
   projection.out = projection_mercator,
   longlat = FALSE,
   out.png = TRUE, 
   add.map = TRUE)
}
plot_figs("chain1")
plot_figs("chain2")
```

